---
title: "A function to play with google drive"
author: "Ronny A. Hernández Mora"
date: "25/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A mockup documented function to connect with the google drive API through R,
check the files that exist in the drive folder, compare to what you have
in your local folder and download just those that you don't have locally.

Could be part of some workflow in an organization that uses google drive as
their site to keep their data files (like a research lab)

```{r, eval = FALSE}
#' @import dplyr
NULL

#' @title Download data from google drive
#' 
#' @author Ronny Alexander Hernández Mora
#' 
#' @description This function will check in a local folder, the existing files
#' and compare which ones are missing from an specific folder in google drive.
#' It will download the missing files
#' 
#' @param drive_path The folder in google drive containing the files
#' @param local_directory The local folder in which we want to download the 
#' files from google drive.
#' 
#' @example 
#' \dontrun{
#' get_drive_data(drive_path = "data_workflows/data", 
#'                local_directory = "datos")
#'}
#'
get_drive_data <- function(drive_path, local_directory) {
  
  options(gargle_oauth_email = "my_email@gmail.com")
  
  # Revisar archivos locales
  archivos_existentes <- fs::dir_ls(local_directory) %>% 
    stringr::str_remove(paste0(local_directory, "/"))
  
  # Revisar archivos en drive
  camino <- drive_path
  
  # Check data available
  archivos_drive <- googledrive::drive_ls(path = camino)
  
  # archivos_drive <- archivos %>% 
  #   select(name)
  
  # Obtener nombres de archivos faltantes
  # Suponiendo que siempre tenemos mas archivos en drive
  archivos_faltantes <- archivos_drive %>% 
    filter(!name %in% archivos_existentes)
  
  # Loop para traerse archivos que estan en drive pero no locales
  for (i in archivos_faltantes$name) {
    archivos_faltantes %>%
      select(id) %>%
      slice(1) %>%
      pull() %>%
      googledrive::drive_download(
        path = paste0("datos/", i), overwrite = TRUE
      )
  }
}
```

After the function is loaded in your Global environment, coupled with a `map()`
and if all the data files have the same variables, we can read everything 
together in just one data frame:

```{r, eval = FALSE}
# Check drive and download data ------------------------------
get_drive_data(drive_path = "data_workflows/data", 
               local_directory = "datos")
# Read data --------------------------------------------------
# Create object with files of interest
files <- dir_ls(path = "datos", glob = "datos/principe_*")

principe <- files %>% 
  map_dfr(read_csv, .id = "file_id")
```


