---
title: "Trying R for remote sensing"
author: "Ronny A. Hernández Mora"
date: "02/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Exporting a COG (Cloud Optimized Geotiff) file from GEE

The idea is to document steps to export from Google Earth Engine a COG file
(or several) to Google Drive to be imported latter in R for analysis. 

## GEE first example

As a first approach, I followed a tutorial from the [Google documentation:](https://developers.google.com/earth-engine/guides/exporting)

```{js, eval = FALSE}
// Load a landsat image and select three bands.
var landsat = ee.Image('LANDSAT/LC08/C01/T1_TOA/LC08_123032_20140515')
  .select(['B4', 'B3', 'B2']);

// Create a geometry representing an export region.
var geometry = ee.Geometry.Rectangle([116.2621, 39.8412, 116.4849, 40.01236]);

// Retrieve the projection information from a band of the original image.
// Call getInfo() on the projection to request a client-side object containing
// the crs and transform information needed for the client-side Export function.
var projection = landsat.select('B2').projection().getInfo();

// Export a cloud-optimized GeoTIFF.
Export.image.toDrive({
  image: landsat,
  description: 'imageToCOGeoTiffExample',
  crs: projection.crs,
  crsTransform: projection.transform,
  region: geometry,
  folder: "GEE_testing",
  fileFormat: 'GeoTIFF',
  formatOptions: {
    cloudOptimized: true
  }
});
```

Then, this file was imported to R with help of the `terra` package:

```{r}
library(terra)

check <- rast("data/imageToCOGeoTiffExample.tif")
```

And, to validate, I plotted the file using a base R function:

```{r}
plot(check)
```



# A failed attempt with raster package

All the code and examples here were created using the information in 
[rspatial.org](https://rspatial.org/raster/rs/1-introduction.html)

## Downloading example data

```{r eval = FALSE}
if (!file.exists('data/rs/samples.rds')) {
    download.file('https://biogeo.ucdavis.edu/data/rspatial/rsdata.zip', dest = 'data/rsdata.zip')
    unzip('data/rsdata.zip', exdir = 'data')
}
```
